---
title: "bluesky"
output: html_document
date: "2022-12-14"
---

## Setup

```{r warning=FALSE, message=FALSE}
library(hdi)
library(tidyverse)
library(rsq)
library(lme4)
library(caret)
```

```{r}
df4 <- read.csv('./cleaned_rhna4_data.csv')
df5 <- read.csv('./cleaned_rhna5_data.csv')

df4 <- select(df4, -X, -MapBlkLot_Master, -mapblklot, -blklot,
              -active, -geometry, -PROPLOC, -RP1PRCLID, -RP1TRACDE)
df5 <- select(df5, -X, -MapBlkLot_Master, -mapblklot, -blklot,
              -active, -geometry, -PROPLOC, -RP1PRCLID, -RP1TRACDE)
```

```{r}
nrow(df4)
nrow(df5)

```

### Remove illogical Values

```{r}
df4$Developed <- df4$Developed > 0
df4 <- df4[df4$ROOMS >= 0,]
df4 <- df4[df4$FBA >= 0,]
df4 <- df4[df4$YRBLT <= 2023,]
summary(df4)
```

```{r}
df5$Developed <- df5$Developed > 0
df5 <- df5[df5$ROOMS >= 0,]
df5 <- df5[df5$FBA >= 0,]
df5 <- df5[df5$YRBLT <= 2023,]
summary(df5)
```

### Exploratory Data Analysis

This is 555 California Street, the fourth largest skyscraper in SF. So the large land value and property values do not seem to be bogus.

```{r}
df4[df4$RP1LNDVAL == max(df4$RP1LNDVAL),]
```

```{r message=FALSE}
ggplot(df4) + 
  geom_histogram(aes(log(1 + ROOMS)))
```

```{r message=FALSE}
ggplot(df4) + 
  geom_histogram(aes(RP1EXMVL1))
```


```{r message=FALSE}
ggplot(df4) + 
  geom_histogram(aes(log(1 + FBA)))
```


```{r message=FALSE}
ggplot(df4) +
  geom_histogram(aes(x=log(RP1LNDVAL + 1))) +
  facet_grid(Developed ~ ., scales = "free_y")
```

```{r message=FALSE}
ggplot(df4) +
  geom_histogram(aes(x=log(RP1IMPVAL + 1))) +
  facet_grid(Developed ~ ., scales = "free_y")
```


```{r}
table(df4$RP1STACDE)
```


### EDA

```{r}
ggplot(df4) +
  geom_point(aes(x=SQFT, y=LAREA), alpha=.1)

```


### Model fitting


I used forward selection to select variables, but it's turned off for creating this R markdown file since it takes an hour to run.

```{r, eval=F}
mod.init <- glm(Developed ~ 1, family='binomial', select(df4, -ZONE))
mod.all <- glm(Developed ~ . -neighborhood, family='binomial', select(df4, -ZONE, -EXEMPTYPE, -RP1PPTVAL, -RP1FXTVAL, -RP1EXMVL2, -RP1CLACDE))
step.mod <- step(mod.init, direction='forward', scope=formula(mod.all))
```

```{r, eval=F}
summary(step.mod)

```

```{r, eval=F}
rsq.kl(step.mod)
```

### Taking log transforms of land value and improvement value helps.

```{r}
mod.log <- glm(Developed ~ general_use_code + district + Upzone_Ratio + 
                 zp_DensRestMulti + zp_OfficeComm + zp_PDRInd + zp_RH3_RM1 + 
                 zp_RH2 + zp_FormBasedMulti + zp_Redev +
                 pipeline2 + pipeline3 + pipeline4 + pipeline5 + pipeline6 + pipeline8 +
                 log1p(RP1LNDVAL) + log1p(RP1IMPVAL) + log1p(ROOMS) + log1p(BEDS) +
                 DEPTH  + Historic + inInventory + Residential_Dummy + 
                 Envelope_1000 + RP1STACDE + log1p(RP1EXMVL1) + log1p(FBA) + YRBLT, 
    family = "binomial", data = df4)

rsq.kl(mod.log)

summary(mod.log)
```

```{r}
caret::confusionMatrix(data=as.factor(ifelse(predict(mod.log, select(df4, -ZONE), type="response") > 0.5, TRUE, FALSE)),  
                       reference=as.factor(df4$Developed))

```

```{r}
caret::confusionMatrix(data=as.factor(as.logical(df4$inInventory)),  
                       reference=as.factor(df4$Developed))

```


```{r}

caret::confusionMatrix(data=as.factor(ifelse(predict(mod.log, 
                                                     select(df5, -ZONE), 
                                                     type="response") > 0.5,
                                             TRUE,
                                             FALSE)),  
                       reference=as.factor(df5$Developed))
```


```{r}

caret::confusionMatrix(data=as.factor(as.logical(df5$inInventory)),
                       reference=as.factor(df5$Developed))

```


```{r message=FALSE}
hist(mod.log$fitted.values)
ggplot(df4, aes(x = mod.log$fitted.values)) +
  geom_histogram() +
  scale_y_log10()

```

# Mixed effects

Turned off temporarily since this is something to evaluate at a later stage
```{r, eval=F}
mod2 <- glmer(Developed ~ . + (1 | RP1CLACDE) - (ZONE + neighborhood + Residential_Dummy + district + RP1CLACDE + CONSTTYPE + EXEMPTYPE + general_use_code), 
              family='binomial', 
              df4)


mod.log.er <- glmer(Developed ~ general_use_code + (1 | district) + Upzone_Ratio + 
                 zp_DensRestMulti + zp_OfficeComm + zp_PDRInd + zp_RH3_RM1 + 
                 zp_RH2 + zp_FormBasedMulti + zp_Redev +
                 pipeline2 + pipeline3 + pipeline4 + pipeline5 + pipeline6 + pipeline8 +
                 log1p(RP1LNDVAL) + log1p(RP1IMPVAL) + log1p(ROOMS) + log1p(BEDS) +
                 DEPTH + Historic + inInventory + Residential_Dummy + 
                 Envelope_1000 + RP1STACDE + log1p(RP1EXMVL1) + log1p(FBA) + YRBLT, 
                 family = "binomial", data = select(df4, -ZONE))

summary(mod.log.er)
```



